---
title: "US 2020 Democratic Candidates"
subtitle: "An Exploration in Campaigns and Net Worth"
author: "Rohan Lewis"
date: "February 22nd, 2020 (Updated May 5th, 2020)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
library(readxl)
library(lubridate)
library(ggplot2)
library(ggrepel)
library(grid)
library(gridExtra)
library(scales)
library(Stack)
library(knitr)
library(lemon)
library(cowplot)

setwd("D:/Data Science/Completed/Democratic Candidates 2020 ")
my_Data <- as.data.frame(read_excel("DC2020.xlsx", sheet = "Data"))
my_Data$Enter <- as.Date(ymd(my_Data$Enter))
my_Data$Exit <- as.Date(ymd(my_Data$Exit))
my_Data = my_Data[rev(order(my_Data$Enter)),]
my_Data$Name = factor(my_Data$Name, levels = unique(my_Data$Name))
rownames(my_Data) <- NULL
```

## Summary

A friend of mine asked me on a Facebook post, "Who is Mike <Bloomberg> ?".  Immediately, I thought he's rich and he started late.  I decided to convey that.

I originally saw something similar [here](https://www.reddit.com/r/dataisbeautiful/comments/f4jwad/net_worth_of_us_2020_democratic_presidential/). 

I did not like it, so I gathered some data.  I found campaign information from [2020 United States presidential election](https://en.wikipedia.org/wiki/2020_United_States_presidential_election#Democratic_Party_nomination) page on Wikipedia.  I found a concise source of net worth for most candidates at [Forbes](https://www.forbes.com/sites/danalexander/2019/08/14/heres-the-net-worth-of-every-2020-presidential-candidate/).

I used R 3.5.3 and ggplot2.

## Campaigns

```{r, fig.width = 10, fig.height = 7}
temp_Data <- my_Data[my_Data$Enter < "2018-01-01",]

g1 <- ggplot(temp_Data, aes(y = Name))
g1 <- g1 + geom_point(aes(x = Enter, color = Campaign), size = 5)
g1 <- g1 + geom_segment(aes(x = Enter, y = Name, xend = as.Date(ymd("2017-12-01")), yend = Name, color = Campaign), size = 1.25)
g1 <- g1 + scale_color_manual(values = c("#009E73", "#E69F00"))
g1 <- g1 + scale_x_date(limits = c(as.Date(ymd("2017-07-16")), as.Date(ymd("2017-12-01"))), date_breaks = "2 months", date_labels = "%B '%y", expand = c(0, 0))
fontcolor <- ifelse(temp_Data$Exit > "2020-02-03", "#009E73", "#E69F00")
fontface <- ifelse(is.na(my_Data$Exit), "bold", "plain")
g1 <- g1 + ggtitle("")
g1 <- g1 + theme(plot.title = element_text(size = 18, face = "bold"),
                 axis.text.x = element_text(hjust = 1, size = 11, angle = 45),
                 axis.text.y = element_text(size = 11, color = fontcolor, face = fontface),
                 axis.title = element_blank(),
                 legend.position = "none")


g2 <- ggplot(my_Data, aes(y = Name))
g2 <- g2 + geom_point(aes(x = Exit, color = Campaign), size = 5)
g2 <- g2 + geom_point(aes(x = Enter, color = Campaign), size = 5, data = my_Data[my_Data$Enter > "2018-01-01",])
g2 <- g2 + geom_segment(aes(x = as.Date(ymd("2018-11-01")), y = Name, xend = Exit, yend = Name, color = Campaign), size = 1.25, data = my_Data[my_Data$Enter < "2018-01-01",])
g2 <- g2 + geom_segment(aes(x = Enter, y = Name, xend = Exit, yend = Name, color = Campaign), size = 1.25, data = my_Data[my_Data$Campaign != "Remaining" & my_Data$Enter > "2018-01-01",])
g2 <- g2 + geom_segment(aes(x = Enter, y = Name, xend = Sys.Date(), yend = Name, color = Campaign), size = 1.5, data = my_Data[my_Data$Campaign == "Remaining",])
g2 <- g2 + scale_color_manual(name = "Campaign Status",
                              values = c("#56B4E9", "#009E73", "#E69F00"),
                              breaks = c("Remaining", "Caucus", "Suspended"),
                              labels = c("Democratic Nominee",
                                         "Suspended after Iowa Caucus",
                                         "Suspended before Iowa Caucus"))
g2 <- g2 + scale_x_date(limits = c(as.Date(ymd("2018-11-01")), Sys.Date()),date_breaks = "3 months", date_labels = "%B '%y", expand = c(0, 0))
g2 <- g2 + scale_y_discrete(position = "right")
g2 <- g2 + xlab("Date of Campaign Announcement and Suspension")
g2 <- g2 + ylab("Candidate")
g2 <- g2 + ggtitle("US 2020 Democratic Candidates")
fontcolor <- ifelse(my_Data$Campaign == "Remaining", "#56B4E9",
                    ifelse(my_Data$Campaign == "Caucus", "#009E73", "#E69F00"))
g2 <- g2 + theme(plot.title = element_text(hjust = -0.2, size = 18, face = "bold"),
                 axis.text.x = element_text(hjust = 1, size = 11, angle = 45),
                 axis.text.y = element_text(size = 11, color = fontcolor, face = fontface),
                 axis.title.x = element_text(hjust = -0.2, size = 14, face = "bold"),
                 axis.title.y = element_text(size = 14, face = "bold"))
legend <- g_legend(g2 + theme(legend.title = element_text(size = 14, face = "bold"),
                              legend.text = element_text(size = 11),
                              legend.position = "right"))
g2 <- g2 + theme(legend.position = "none")

grid.arrange(g1, g2, legend, layout_matrix = matrix(rbind(c(1, 2, 2, 2),
                                                          c(1, 2, 2, 2),
                                                          c(NA, 2, 2, 2),
                                                          c(NA, 2, 2, 2),
                                                          c(3, 2, 2, 2),
                                                          c(NA, 2, 2, 2),
                                                          c(NA, 2, 2, 2),
                                                          c(NA, 2, 2, 2)),
                                                    ncol = 4))
```

## Note

From here, I modified the dataset.

1) I eliminated Gravel, Hickenlooper, Messam, Ojeda, Patrick, and Salwell. I did not have Net Worth Data for them.

2) I eliminated all Suspensions, and only looked at Announcements.

```{r,}
new_Data <- my_Data[!is.na(my_Data$Worth),]
new_Data$Name <- as.factor(as.character(new_Data$Name))
rownames(new_Data) <- NULL
```


## Net Worth

```{r, fig.width = 10, fig.height = 6.5}
options(scipen = 999)
temp <- new_Data[new_Data$Worth < 3000000,]
g3 <- ggplot(temp, aes(x = Enter, y = Worth))
g3 <- g3 + scale_color_manual(values = c("#009E73", "#E69F00", "#56B4E9"))
g3 <- g3 + geom_point(aes(size = Worth, color = Campaign))
g3 <- g3 + geom_text_repel(aes(label = Name), color = "black", size = 5)
g3 <- g3 + scale_size_continuous(range = c(50 / sqrt(max(temp$Worth) / min(temp$Worth)), 50))
g3 <- g3 + scale_x_date("", limits = c(as.Date(ymd("2017-10-15")), as.Date(ymd("2017-11-15"))), date_breaks = "1 month", date_labels = "%B '%y")
g3 <- g3 + scale_y_continuous("Net Worth as of August '19", label = dollar, limits = c(0, 3000000))
g3 <- g3 + ggtitle("")
g3 <- g3 + theme(plot.title = element_text(hjust = 0.5, face = "bold"),
               axis.text.x = element_text(hjust = 1, size = 12, angle = 45),
               axis.text.y = element_text(size = 12),
               axis.title = element_text(size = 14, face = "bold"),
               legend.position = "none")

g4 <- ggplot(temp, aes(x = Enter, y = Worth))
g4 <- g4 + scale_color_manual(values = c("#009E73", "#E69F00", "#56B4E9"))
g4 <- g4 + geom_point(aes(size = Worth, color = Campaign))
g4 <- g4 + geom_text_repel(aes(label = Name), color = "black", size = 5)
g4 <- g4 + scale_size_continuous(range = c(50 / sqrt(max(temp$Worth) / min(temp$Worth)), 50))
g4 <- g4 + scale_x_date(limits = c(as.Date(ymd("2019-01-01")), as.Date(ymd("2019-06-01"))), "Date of Campaign Announcement", date_breaks = "1 month", date_labels = "%B '%y")
g4 <- g4 + scale_y_continuous("", labels = NULL, limits = c(0, 3000000))
g4 <- g4 + ggtitle("US 2020 Democratic Candidates")
g4 <- g4 + theme(plot.title = element_text(hjust = 0, face = "bold"),
               axis.text.x = element_text(hjust = 1, size = 12, angle = 45),
               axis.text.y = element_text(size = 12),
               axis.title = element_text(hjust = 0, size = 14, face = "bold"),
               legend.position = "none")

grid.arrange(g3, g4, layout_matrix = matrix(c(1, 1, 2, 2, 2, 2, 2), ncol = 7))





temp <- new_Data[new_Data$Worth < 20000000,]
g5 <- ggplot(temp, aes(x = Enter, y = Worth))
g5 <- g5 + scale_color_manual(values = c("#009E73", "#56B4E9", "#E69F00"))
g5 <- g5 + geom_point(aes(size = Worth, color = Campaign))
g5 <- g5 + geom_text_repel(data = subset(temp, Worth > 2000000), aes(label = Name), color = "black", size = 5)
g5 <- g5 + scale_size_continuous(range = c(50 / sqrt(max(temp$Worth) / min(temp$Worth)), 50))
g5 <- g5 + scale_x_date("", limits = c(as.Date(ymd("2017-10-15")), as.Date(ymd("2017-11-15"))), date_breaks = "1 month", date_labels = "%B '%y")
g5 <- g5 + scale_y_continuous("Net Worth as of August '19", label = dollar, limits = c(0, 20000000))
g5 <- g5 + ggtitle("")
g5 <- g5 + theme(plot.title = element_text(hjust = 0.5, face = "bold"),
                 axis.text.x = element_text(hjust = 1, size = 12, angle = 45),
                 axis.text.y = element_text(size = 12),
                 axis.title = element_text(size = 14, face = "bold"),
                 legend.position = "none")

g6 <- ggplot(temp, aes(x = Enter, y = Worth))
g6 <- g6 + scale_color_manual(values = c("#009E73", "#56B4E9", "#E69F00"))
g6 <- g6 + geom_point(aes(size = Worth, color = Campaign))
g6 <- g6 + geom_text_repel(data = subset(temp, Worth > 2000000), aes(label = Name), color = "black", size = 5)
g6 <- g6 + scale_size_continuous(range = c(50 / sqrt(max(temp$Worth) / min(temp$Worth)), 50))
g6 <- g6 + scale_x_date(limits = c(as.Date(ymd("2019-01-01")), as.Date(ymd("2019-07-01"))), "Date of Campaign Announcement", date_breaks = "1 month", date_labels = "%B '%y")
g6 <- g6 + scale_y_continuous("", labels = NULL, limits = c(0, 20000000))
g6 <- g6 + ggtitle("US 2020 Democratic Candidates")
g6 <- g6 + theme(plot.title = element_text(hjust = 0, face = "bold"),
                 axis.text.x = element_text(hjust = 1, size = 12, angle = 45),
                 axis.text.y = element_text(size = 12),
                 axis.title = element_text(hjust = 0, size = 14, face = "bold"),
                 legend.position = "none")

grid.arrange(g5, g6, layout_matrix = matrix(c(1, 1, 2, 2, 2, 2, 2), ncol = 7))





temp <- new_Data[new_Data$Worth < 250000000,]
g7 <- ggplot(temp, aes(x = Enter, y = Worth))
g7 <- g7 + scale_color_manual(values = c("#009E73", "#56B4E9", "#E69F00"))
g7 <- g7 + geom_point(aes(size = Worth, color = Campaign))
g7 <- g7 + geom_text_repel(data = subset(temp, Worth > 14000000), aes(label = Name), color = "black", size = 5)
g7 <- g7 + scale_size_continuous(range = c(50 / sqrt(max(temp$Worth) / min(temp$Worth)), 50))
g7 <- g7 + scale_x_date("", limits = c(as.Date(ymd("2017-06-01")), as.Date(ymd("2017-12-01"))), date_breaks = "3 months", date_labels = "%B '%y")
g7 <- g7 + scale_y_continuous("Net Worth as of August '19", label = dollar, limits = c(0, 250000000))
g7 <- g7 + ggtitle("")
g7 <- g7 + theme(plot.title = element_text(hjust = 0.5, face = "bold"),
                 axis.text.x = element_text(hjust = 1, size = 12, angle = 45),
                 axis.text.y = element_text(size = 12),
                 axis.title = element_text(size = 14, face = "bold"),
                 legend.position = "none")

g8 <- ggplot(temp, aes(x = Enter, y = Worth))
g8 <- g8 + scale_color_manual(values = c("#009E73", "#56B4E9", "#E69F00"))
g8 <- g8 + geom_point(aes(size = Worth, color = Campaign))
g8 <- g8 + geom_text_repel(data = subset(temp, Worth > 14000000), aes(label = Name), color = "black", size = 5)
g8 <- g8 + scale_size_continuous(range = c(50 / sqrt(max(temp$Worth) / min(temp$Worth)), 50))
g8 <- g8 + scale_x_date(limits = c(as.Date(ymd("2019-01-01")), as.Date(ymd("2019-07-01"))), "Date of Campaign Announcement", date_breaks = "1 month", date_labels = "%B '%y")
g8 <- g8 + scale_y_continuous("", labels = NULL, limits = c(0, 250000000))
g8 <- g8 + ggtitle("US 2020 Democratic Candidates")
g8 <- g8 + theme(plot.title = element_text(hjust = 0, face = "bold"),
                 axis.text.x = element_text(hjust = 1, size = 12, angle = 45),
                 axis.text.y = element_text(size = 12),
                 axis.title = element_text(hjust = 0, size = 14, face = "bold"),
                 legend.position = "none")

grid.arrange(g7, g8, layout_matrix = matrix(c(1, 1, 2, 2, 2, 2, 2), ncol = 7))





temp <- new_Data[new_Data$Worth < 2000000000,]
g9 <- ggplot(temp, aes(x = Enter, y = Worth))
g9 <- g9 + scale_color_manual(values = c("#009E73", "#56B4E9", "#E69F00"))
g9 <- g9 + geom_point(aes(size = Worth, color = Campaign))
g9 <- g9 + geom_text_repel(data = subset(temp, Worth > 100000000), aes(label = Name), color = "black", size = 5)
g9 <- g9 + scale_size_continuous(range = c(50 / sqrt(max(temp$Worth) / min(temp$Worth)), 50))
g9 <- g9 + scale_x_date("", limits = c(as.Date(ymd("2017-06-01")), as.Date(ymd("2017-12-01"))), date_breaks = "3 months", date_labels = "%B '%y")
g9 <- g9 + scale_y_continuous("Net Worth as of August '19", label = dollar, limits = c(0, 2000000000))
g9 <- g9 + ggtitle("")
g9 <- g9 + theme(plot.title = element_text(hjust = 0.5, face = "bold"),
                 axis.text.x = element_text(hjust = 1, size = 12, angle = 45),
                 axis.text.y = element_text(size = 12),
                 axis.title = element_text(size = 14, face = "bold"),
                 legend.position = "none")

g10 <- ggplot(temp, aes(x = Enter, y = Worth))
g10 <- g10 + scale_color_manual(values = c("#009E73", "#56B4E9", "#E69F00"))
g10 <- g10 + geom_point(aes(size = Worth, color = Campaign))
g10 <- g10 + geom_text_repel(data = subset(temp, Worth > 100000000), aes(label = Name), color = "black", size = 5)
g10 <- g10 + scale_size_continuous(range = c(50 / sqrt(max(temp$Worth) / min(temp$Worth)), 50))
g10 <- g10 + scale_x_date(limits = c(as.Date(ymd("2019-01-01")), as.Date(ymd("2019-08-01"))), "Date of Campaign Announcement", date_breaks = "1 month", date_labels = "%B '%y")
g10 <- g10 + scale_y_continuous("", labels = NULL, limits = c(0, 2000000000))
g10 <- g10 + ggtitle("US 2020 Democratic Candidates")
g10 <- g10 + theme(plot.title = element_text(hjust = 0, face = "bold"),
                 axis.text.x = element_text(hjust = 1, size = 12, angle = 45),
                 axis.text.y = element_text(size = 12),
                 axis.title = element_text(hjust = 0, size = 14, face = "bold"),
                 legend.position = "none")

grid.arrange(g9, g10, layout_matrix = matrix(c(1, 1, 2, 2, 2, 2, 2), ncol = 7))



temp <- new_Data[new_Data$Worth < 70000000000,]
g11 <- ggplot(temp, aes(x = Enter, y = Worth))
g11 <- g11 + scale_color_manual(values = c("#009E73", "#56B4E9", "#E69F00"))
g11 <- g11 + geom_point(aes(size = Worth, color = Campaign))
g11 <- g11 + geom_text_repel(data = subset(temp, Worth > 1000000000), aes(label = Name), color = "black", size = 5)
g11 <- g11 + scale_size_continuous(range = c(50 / sqrt(max(temp$Worth) / min(temp$Worth)), 50))
g11 <- g11 + scale_x_date("", limits = c(as.Date(ymd("2017-06-01")), as.Date(ymd("2017-12-01"))), date_breaks = "3 months", date_labels = "%B '%y")
g11 <- g11 + scale_y_continuous("Net Worth as of August '19", label = dollar, limits = c(0, 70000000000))
g11 <- g11 + ggtitle("")
g11 <- g11 + theme(plot.title = element_text(hjust = 0.5, face = "bold"),
                 axis.text.x = element_text(hjust = 1, size = 12, angle = 45),
                 axis.text.y = element_text(size = 12),
                 axis.title = element_text(size = 14, face = "bold"),
                 legend.position = "none")

g12 <- ggplot(temp, aes(x = Enter, y = Worth))
g12 <- g12 + scale_color_manual(values = c("#009E73", "#56B4E9", "#E69F00"))
g12 <- g12 + geom_point(aes(size = Worth, color = Campaign))
g12 <- g12 + geom_text_repel(data = subset(temp, Worth > 1000000000), aes(label = Name), color = "black", size = 5)
g12 <- g12 + scale_size_continuous(range = c(50 / sqrt(max(temp$Worth) / min(temp$Worth)), 50))
g12 <- g12 + scale_x_date(limits = c(as.Date(ymd("2019-01-01")), as.Date(ymd("2019-12-15"))), "Date of Campaign Announcement", date_breaks = "1 month", date_labels = "%B '%y")
g12 <- g12 + scale_y_continuous("", labels = NULL, limits = c(0, 70000000000))
g12 <- g12 + ggtitle("US 2020 Democratic Candidates")
g12 <- g12 + theme(plot.title = element_text(hjust = 0, face = "bold"),
                 axis.text.x = element_text(hjust = 1, size = 12, angle = 45),
                 axis.text.y = element_text(size = 12),
                 axis.title = element_text(hjust = 0, size = 14, face = "bold"),
                 legend.position = "none")

grid.arrange(g11, g12, layout_matrix = matrix(c(1, 1, 2, 2, 2, 2, 2), ncol = 7))
```

***

# Appendix 
The complete code and RMarkdown can be found at my  [GitHub](https://github.com/HumanRickshaw/US-2020-Democratic-Candidates.git).


## Summary
```{r}
kable(my_Data)
```

## Note
```{r}
kable(new_Data)
```